\section{Acknowledgements}

After defending my PhD thesis in 2020, I swore to take a good break
from the stress of academia, and spent a gap year in Athens, Greece.
This turned out to be the most prolific year of my career. In the absence
of external pressure, I could finally get some research done.

It was during that summer of 2021 that I started designing a new course on
\emph{Blockchain Foundations}. This 56-hour course aimed to answer
three questions:

\begin{enumerate}
  \item What are blockchains?
  \item How do they work?
  \item Why are they secure?
\end{enumerate}

I found that there was already a corpus of works describing the
details of blockchains. However, that literature was not quite what
I was looking for. Some of it was scattered across informal blog posts
that explained high-level ideas tailored towards end users or
hobby scientists, and never went down to the mathematical details.
The writing quality here is generally also poor.
Other writings were \emph{very} precise on the engineering level
--- talking about this or that byte of a packet --- but never explained
the \emph{why} behind the design decisions, and in particular they
were missing security proofs against arbitrary adversaries.
Lastly, a body of works of high quality \emph{does} provide mathematical
proofs and justification
for the \emph{whys}, but these are in the form of scientific papers
that are extremely dense and beyond reach of even advanced graduate
students, let alone undergraduates dipping their feet on blockchain
science for the first time. I decided it's about time to write a
series of lecture notes on the foundations to accompany the lectures.

As I was designing the course, I trialed it to a group of my computer
science colleagues with no prior blockchain experience:
Giannis Gkoulioumis, Nikolaos Kamarinakis, Apostolos Tzinas.
Over the course of that summer, we spent 52-hours together over notebooks
of notes, discussing the proofs of \emph{bitcoin backbone} and the
construction of Merkle trees. They also solved a series of programming
exercises which I designed to accompany the course and pertained to
creating their own blockchain from scratch, each building their own
full node. Their feedback gave me the opportunity to refine the course
and prepare it for teaching.

I had the first opportunity to teach this course --- which I nicknamed
\emph{Marabu} after the eponymous poem of Nikos Kavvadias --- during the
spring quarter of 2022 at Stanford University. It was a graduate-level course
named EE374 - Blockchain Foundations. While most of the material was
based on the past summer, my co-instructor David Tse and I redesigned
parts of the course to meet the time constraints of the quarter system,
and to include some new material. To our surprise, the course was mostly
attended by undergraduates.
The course was well received,
largerly due to great work of our teaching assistants, Srivatsan Sridhar
and Kamilla Nazirkhanova.

At the time, the lecture notes well still
in my handwritten notebooks. As I taught the lectures on the whiteboard,
several students helped out with the digitizing the lecture notes to
form the first version of this book. These scribes were
Kaylee Renae George, Kenan Hasanaliyev, Koren Gilbai,
Ben Choi, Stephen Su, Nathaniel Masfen-Yan,
Schwinn Saereesitthipitak, Kachachan Chotitamnavee, Alan Zhang,
Taher Poonawala, Scott Hickmann,
Lyron Co Ting Keh, Sam Liokumovich,
Kaili Wang, Andrej Elez,
Edward Vendrow, Cathy Zhou,
Yifan Yang,
Michael Nath, Coleman Smith,
Solal Afota,
Suppakit Waiwitlikhit, Lora Xie,
Bryan Chiang, Lucas Xia,
% below here are for David's notes, but let's keep their names
Albert Pun, Gordon Chi,
Jack Liu,
Neetish Sharma, Sergio Charles,
Priyanka Mathikshara, and John Guibas.
Even though these notes were later rewritten many times,
I'm deeply grateful to all, because they made the first version happen,
and it would not have started without them.

We repeated this course during winter quarter of 2023.
There, our teaching assistants were Kenan Hasanaliyev
and Scott Hickmann who further helped refine the course material and notes.
